### 2021-11-06 plos_latex_template.tex
rj and ast resolving lots of comments

%andrea: matt's original sentence: reaching users and developers of the R language from more than 120 countries.

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break. need some ligation, something like "Despite the many efforts in this direction, ..."}
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members, all of which are readily provided to non-marginalized people \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, \as{prejudice against disable people ( ableism)}, among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021}, with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article focuses on practices for hosting inclusive conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the past year of organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}

@@

% Dorothea: I like the intro as it is now, and I don't think the following paragraph is necessary.
% Rocío: I'm leaving this out now and seeing if we don't need this description of communities
%% Rocío: Main message of the paragraph: We do this for communities and people. What do we mean by communities and why are they important? (Why are they important may need some text)
%This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
%Some scholarly or technological communities are formally structured as learned societies (e.g. the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g. R User Groups or Python User Groups), and others may not have any established structure. 
%For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.

% keeping references:
%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018, marks2021meeting}.

@@

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. \fede{General thought: I like the way we say WE and YOU. I still think I saw most such 10SR papers speak to the readers addressing them with the roles. Which would be "Organizers", "Attendees". Just my 2 cents as avid reader of 10SR papers ;)}
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
Will a more diverse conference translate into a gender distribution of your speakers that is representative of the general population? 
Would it be the presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
The answer to these questions depend on your field, region, or community.

%batool: There are a few references which define privilege and might be good to add here including: Friedman, S., O’Brien, D., & McDonald, I. (2021). Deflecting Privilege: Class Identity and the Intergenerational Self. Sociology, 55(4), 716–733. https://doi.org/10.1177/0038038520982225
%batool: Another one is: Crevani, L. (2019). Privilege in place: How organisational practices contribute to meshing privilege in place. Scandinavian Journal of Management, 35(2). https://doi.org/10.1016/j.scaman.2018.09.002
%batool: please note that I didn't read both entire papers I mentioned above.
%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).
% Rocío: I feel like we don't need to cite any reference about privilege the way it is now. But feel free to disagree.
%Dorothea: I agree with Rocío. citations for priviledge feel like citation for gravity...the current wording is not saying "there is priviledge, as shown by xyz. it says there is priviledge and we need to see where it is particularly strong in our field.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a Code of Conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly. 
Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive, victim-centered manner. It may also be more understanding of different disciplinary cultures and geographical and cultural considerations. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.
% Rocío: two of these sentences, diverse CoC team, and communication of CoC, may belong to rules organizing team and communication, respectively.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
When working with people from marginalized groups do not assume that self-nomination and voting will work as mechanisms to choose leadership positions. Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.

% Saranjeet: At times regional institutes and universities do not support nor recognize the time put in organization tasks. As such one needs to carry out their day work for the institute while working extra hours for organizing. It might be challenging to find a way to recognize organizers that works internationally. However, there should certainly be some mechanism put into place which recognizes the time of the organizers.
In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

@@

\section{Consciously counteract bias in your spotlight roles}
% in the conference program
\label{rule_unbias}

@@

Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes. 
Selection committees need to be reminded of such biases, and try to counteract them when evaluating submissions and presentations, which should have clear evaluation criteria.
\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success

Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}. Research about these topics is equally undervalued compared to other areas of knowledge production. 
Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
%Defy the stereotypical criteria for merit by acknowledging these community practices and the people behind them.
You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones or deciding to rotate the favored timezone every year, without depending on a conference central location.

@@

% Saranjeet: Not to include in this paper, but a general suggestion. Why not select a platform which helps in live broadcasting of the conference on LinkedIn as well? For example, Streamyard (https://streamyard.com/) is one such multistreaming platform, that can simultaneously stream to YouTube, LinkedIn, Twitter, and other platforms

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
Alternatively, a conference could have a hybrid format with an in-person and an online components \la{that articulate well into a hybrid format that maximizes the experience of everyone attending your conference}. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.

In fact, experts in theorizing about and building communities of practice [Wenger, 1999] who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future. These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences. These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020]. Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups. The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space. 
%an online component is not inclusive \textit{per se} without taking many other precautions

%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing and 
some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}). 
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community. 
It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference %LA cita 
.
Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}

% we need to check if something here (newbies!) is not missing in the above paragraphs
% As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)

@@

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.

@@

Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone %ast: refs here.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communication channels managed by the conference organizers (see \textbf{Rule \ref{rule_communication}}).

@@

% Not citing ninerBetterWhomLeveling2021 anymore for now.

@@

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
to emphasize that they* are particularly welcome in this event. This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}). Try to come up with creative ways to show that everyone is seen, respected, and welcome. \la{For example, useR! 2021 created "Margot, the marmot" a mascot for the conference. Margot is a cartoon with different designs reflecting folks that the conference wanted to reach. In particular, Margot wears different scarfs. The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc. (Figure 2)."}.
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
% allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.

@@

% Rocío: main idea: Budgets are limited and need to choose priorities.
Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering) % Rocío: can we give examples?
% Dorothea: this depends so much on the conference type. I wanted to write "venue"but this is not true of course, so I don't know what to add. %ast i added a couple more, i meant the most "obvious" budget items  in contrast to the less obvious
allocation of resources has to be intentional to support the goals of inclusion of the conference. 
It is important to estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
Additional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). 
Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion, than making generic demands for funding.
%Dorothea: I slightly disagree with that one. Maybe just leave the first part out?

>>>>

@@

% Rocío: Main message of the paragraph: Why we are writing this piece
Conferences, from the Latin \textit{conferre}--`to bring together', are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.\fede{between this and the next sentence, there is a strong break. need some ligation, something like "Despite the many efforts in this direction, ..."}
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. 
Lack of representation and unwelcoming--or overtly aggressive--environments exclude people from opportunities for learning, sharing, and producing knowledge, and becoming active community members \cite{hendersonThoughtfulGatheringsGendering2020}.
Exclusionary conference experiences can divert career paths, affect lives, and drive people out of academia \cite{biggsAcademicConferenceChilly2018, hendersonThoughtfulGatheringsGendering2020}. 
This message is not new. 
Barriers such as unattainable registration costs, sexism, prejudice against disable people (ableism), among others, have already been exposed and discussed in the literature \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}, 
and some proposals for more inclusive conferences have been put into practice \cite{gichoraTenSimpleRules2010a, levitisCenteringInclusivityDesign2021, atkinsonJournalMedicine20202021, foramittiVirtuesVirtualConferences2021, ninerBetterWhomLeveling2021, rabyMovingAcademicConferences2021, noauthor_discover2021},
with a primary focus on the online format to open doors for inclusion.\fede{do we need maybe a short definition of the concepts of diversity-equity-inclusion for us/refer to something?} %LA: https://www.cscce.org/resources/organizing-community-events/

% Rocío: Main message of the paragraph: Roadmap of the paper
This article focuses on practices for hosting inclusive conferences.
The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
Here, we share the lessons learned within the time organizing useR! 2021, summarized as ten simple rules towards hosting an inclusive conference.
The rules are organized in three groups (Figure \ref{fig:diagram}). %original source: https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing
%sara: still need to better format figure size and get the alt text right :P
Group 1 includes rules 1, 2, and 10, which refer to pillars of an inclusive conference: embracing diversity in all its dimensions, creating a safe and welcoming environment for everyone, and making the conference part of a long-term process for inclusion.
The second group includes rules 3 and 4 and focuses on the people who participate in the conference. 
Rule 3 refers to the importance of working with an inclusive and diverse organizing team, and Rule 4 concerns the necessity of counteracting implicit and systemic bias from spotlight roles like keynote speakers, other presenters, program committee members, or other session chairs. 
The third group includes rules 5 to 9. These rules are about components of the conference that should be carefully planned for: an online component, accessibility for people with disabilities, language inclusiveness, a welcoming communication strategy, and financial resources to support inclusion. 
These rules apply equally to conferences where the expected audience is academic, from the industry, or a mixed group. While these rules are inspired in a global virtual experience they also apply at the regional or local level and can inform in-person and hybrid events.
\fede{maybe something like: "Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes"}

@@

@@

% Rocío: more effort towards including the more excluded
Recognizing our privileges--unearned advantages given by society to some people but not all--particularly in our field and in our scientific or professional community, will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include.
Investing more effort towards the most excluded groups does not mean neglecting the others, but it will guide the vision of diversity for your conference--and your strategies to achieve it.
This may translate into different actions depending on your field, region, or community, e.g. a gender distribution of your speakers that is representative of the general population, presence of racialized people --especially Black people-- among the head organizers, speakers, and attendees, LGBTQIA+ friendly-spaces, or community participation from key geographic regions.

@@

% Rocío: main idea: Code of conduct for a safe place
Adopting a Code of Conduct and gathering a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
The Code of Conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The Code of Conduct team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize accordingly creating protocols on how to respond. This should be done during the early phases of the conference planning because Code of conduct violations can happen before the conference (e.g. social media, organization meetings).  
Since people from marginalized groups are more likely to be victims of Code of Conduct violations and to see their claims dismissed due to power dynamics, assembling a diverse Code of Conduct team will be essential to deal with discrimination and harassment issues and the power dynamics underneath them and deal with them in a sensitive manner. It may also be more understanding of different disciplinary cultures and geographical and cultural considerations. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
When working with people from marginalized groups do not assume that self-nomination and voting will work as mechanisms to choose leadership positions. Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.  
% Saranjeet: At times regional institutes and universities do not support nor recognize the time put in organization tasks. As such one needs to carry out their day work for the institute while working extra hours for organizing. It might be challenging to find a way to recognize organizers that works internationally. However, there should certainly be some mechanism put into place which recognizes the time of the organizers.
% Rocío: Add other types of recognition besides money (website, public recognition on social media, during the conference, etc.)
In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

@@

\section{Consciously counteract bias in the conference program}
\label{rule_unbias}

@@

% our committees are biased and unconscious bias exists
Having a chance to present at a conference is an opportunity to gain--or keep--visibility in the community or field. 
Even diverse selection committees can be subjected to unconscious bias, by interpreting information such as the location of the institution or the names of the authors and attaching value to their work according to negative or positive stereotypes. 
Selection committees need to be reminded of such biases, and try to counteract them when evaluating submissions and presentations, which should have clear evaluation criteria.
\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
Activities such as scientific publication and development of software are often regarded as higher value than care-taking roles like community building or teaching. 
These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}. Research about these topics is equally undervalued compared to other areas of knowledge production. 
Counteracting bias also means broadening the range of topics of the conference and giving spotlight/visibility to usually undervalued research and the people behind it.
%Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
You can also reframe the awards ceremony to acknowledge those who contributed to community building, presented teaching materials, and assumed other care-taking roles.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction at conferences is priceless but often this is true and possible only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive because they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of attendees but in terms of speakers  \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several time zones or deciding to rotate the favored time zone every year, without depending on a conference central location.

@@

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing and some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues). 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012, blackEngenderingBelongingThoughtful2020}). 
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that no single networking activity will ever serve the whole community. 
It is worth trying new and varied activities centered around different groups of people outside the mainstream of your conference.
Some ideas in this line are: offering the option of written chat instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.\fede{another activity: virtual city tours, I had it as a corollary activity for a summer school} \la{Some more specifics on how to achieve these suggestions in this paper: https://osf.io/k3bfn/}
% we need to check if something here (newbies!) is not missing in the above paragraphs
% As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above \cite{jooKeepOnlineOption2021}.
Alternatively, a conference could have a hybrid format with an in-person and an online components that articulate well and maximize the experience of everyone attending your conference. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{jooKeepOnlineOption2021,ninerBetterWhomLeveling2021}.

In fact, experts in theorizing about and building communities of practice [Wenger, 1999] who have been holding hybrid events since 2011 [Wenger-Tayner, 2021] consider that inclusive hybrid events are the future. These authors emphasize the need that all participants of hybrid events must go through the same event quality to the point that in-person and online participants pay the same registration amount to participate in high-quality hybrid conferences. These authors recommend having a buddy system, where every person attending online has an in-person attendant that makes sure the online person can follow the event and participate in the event as if they were in the room [Wenger-Trayner, 2020]. Wenger-Trayner (2020, 2021) also recommends having a chat function for back-channel conversations between all attendants, a virtual space for all online attendants to talk to each other, and allocating time to small group conversations, often mixing in-person and online groups. The physical venue will have to include multiple ways to connect with online folks such as tablets on the meeting venue tables and cameras that allow online attendants to follow who is speaking in the in-person space.

% [Wenger, 1999] https://doi.org/10.1017/CBO9780511803932
% [Wenger-Trayner, 2020] https://wenger-trayner.com/all/in-person-and-online-events/
% [Wenger-Trayner, 2021] https://wenger-trayner.com/reflections/in-person-online-hybrid-the-future/

@@

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where participation in the organizing team allows disabled people to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it are not substitutes for real-life experience \cite{costanzachockDesign2020}. Planning for accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct. 
For in-person conferences, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities (e.g., dyslexia, anxiety, autism and tis spectrum). This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, provide chairs in open spaces, and menus that menus that account for the dietary requirements participants need to respect.

@@

Accessibility practices should also include social events and networking, and include activities that do not restrict participation based on body type or ability. Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communication channels  (see \textbf{Rule \ref{rule_communication}}).

@@

@@

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
Include all the work to make the conference a safe and inclusive environment in your communication strategy, 
to emphasize that they* are particularly welcome in this event. This could be done through publishing a diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the Code of Conduct, the accessibility guidelines, and advertising financial support (see Rule \ref{rule_financial}). Try to come up with creative ways to show that everyone is seen, respected, and welcome. \la{For example, useR! 2021 created "Margot, the marmot" a mascot for the conference. Margot is a cartoon with different designs reflecting folks that the conference wanted to reach. In particular, Margot wears different scarfs. The scarfs include colors of flags from different communities in R such as MiR´s, Rainbow R, ..., etc. (Figure 2)."}.
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists)--
part of your community may use a different platform than the ones adopted by people who regularly attend the conference. 
% rj: allude to the diverse organizing team - to tap into groups, contexts and make choices on platforms and ways to communicate
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too.

@@

% Rocío: main idea: Budgets are limited and need to choose priorities.
Conference budgets are limited and rely mostly on sponsorship and attendance fees. 
While some expenses might be more or less fixed or known, (e.g., venue, internet platforms, catering), allocation of resources has to be intentional to support the goals of inclusion of the conference. 
It is important to estimate the costs for these practices and define your priorities in advance (e.g. paying the organizing team, Code of Conduct training, captioning).
Additional support for attendees should also be considered in the budget: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). 
Consider that an online conference might reduce organization costs (e.g. no rental costs for a physical venue), allowing to redirect the money towards other inclusive priorities. 
When asking for sponsorship, it might be easier to justify supporting concrete actions towards inclusion than making generic demands for funding.
